---
title: "PV data - Parameter estimation and sample generation"
author: "Martina Kuchlbauer"
date: "Tuesday, May 17, 2016"
output:
  pdf_document:
    number_sections: yes
---


```{r, include=FALSE,cache=FALSE}
PVdata <- read.csv("~/Uni/8.Semester/CaseStudies/CaseStudy/data/PVdata2csv.csv", header=FALSE) #load data
#PVdata <- read.csv("~/Documents/CaseStudy/data/PVdata2csv.csv", header=FALSE) #load data
library(mvnmle) #package for parameter estimation
library(MASS)
library(fitdistrplus)
```

We consider the data given in PVdata2.csv.
Let's generate a matrix where each line represents a day and each column represents one minute of this day: 

```{r}
PV1<-PVdata[1:1440,1] #24*60=1440
for (i in 1:30) {
  PV1<-cbind(PV1,PVdata[((i*1440)+1):((i+1)*1440),1])
}
for(j in 2:12){
  for(i in 1:31){
    PV1<-cbind(PV1,PVdata[(((i-1)*1440)+1):(i*1440),j])
  }
}
PV1<-t(PV1)
```


#Normal distribution

##Parameter estimation

###Parameter estimation - multivariate for 1h intervals\

We estimate the values of expectation and the covariance matrix under the assumption of a **multivariate** normal distribution for intervals of 1h: 

```{r}
PV1h<-matrix(rep(0,8928),nrow=372) #hourly values -> take means, 24*372=8928
for(i in 1:372){
  for (j in 1:24){
    PV1h[i,j]<-mean(PV1[i,((j-1)*60+1):(j*60)])
  }
}

estimates_n_dep<-mlest(PV1h) #under assumption of no independence: hourly means and covariance matrix

#estimates_n_dep$muhat #estimate of mean (mu)
#estimates_n_dep$sigmahat #estimate of covariance matrix (sigma)
```

Let's visualize the expected value we estimated: 

```{r}
plot(estimates_n_dep$muhat,xlab="time [h]", ylab="radiation", type="l", col="red", ylim=c(0,3000))
legend(17,3000, c("exp value"), col=c("red"), lty=c(1))
```


###Parameter estimation - independent for 15min intervals\


Using the package above it is not possible to analyse a multivariate normal distribution with more than 50 variables, hence we continue the analysis with the assumption of $4\cdot 24$ **independently** distributed variables for every 15 minutes. 


```{r}
PV1quh<-matrix(rep(0,(8928*4)),nrow=372) #quarter hourly values -> take means, 24*372=8928
for(i in 1:372){
  for (j in 1:(24*4)){
    PV1quh[i,j]<-mean(PV1[i,((j-1)*15+1):(j*15)])
  }
}

estimates_n_ind<-matrix(rep(0,2*96),nrow=2) #under assumption of independence, 4*24=96=T
for (i in 1:(24*4)){
  estimates_n_ind[1,i]<-fitdistr(PV1quh[,i],"normal")$estimate[1] #estimate of mean (mu)
  estimates_n_ind[2,i]<-fitdistr(PV1quh[,i],"normal")$estimate[2] #estimate of std error (sigma)
  }
#estimates_n_ind

```

Let's visualize the expected value and standard errors we estimated: 

```{r}
plot(estimates_n_ind[1,],xlab="time [15min]", ylab="radiation", type="l", col="red", ylim=c(0,3500))
lines(estimates_n_ind[1,]+estimates_n_ind[2,],type="l")
lines(estimates_n_ind[1,]-estimates_n_ind[2,],type="l")
legend(65,3400, c("exp value", "+/- std error"), col=c("red","black"), lty=c(1,1))

```

###Parameter estimation - sum of RVs 

We now estimate the parameters under the assumption that the radiation values are distributed according to $\frac{1}{4}\cdot (X+Y+Z+W)$ where $X,Y,Z$ and $W$ are normally distributed.\ 

Let $X$ be distributed with the parameters we estimated for the **independently** normal distribution for intervals of 15min.\

For every interval of 1h, let $Y$ be the random variable distributed according to a univariate normal distribution. 

```{r}
estimates_n_sum2<-matrix(rep(0,2*24),nrow=2) #under assumption of independence
for (i in 1:(24)){
  estimates_n_sum2[1,i]<-fitdistr(PV1h[,i],"normal")$estimate[1] #estimate of mean (mu)
  estimates_n_sum2[2,i]<-fitdistr(PV1h[,i],"normal")$estimate[2] #estimate of std error (sigma)
  }
```

For every interval of 3h, let $Z$ be the random variable distributed according to a univariate normal distribution. 

```{r}
PV13h<-matrix(rep(0,(8*372)),nrow=372) #3h values -> take means, 8*372=2976
for(i in 1:372){
  for (j in 1:(8)){
    PV13h[i,j]<-mean(PV1h[i,((j-1)*3+1):(j*3)])
  }
}
estimates_n_sum3<-matrix(rep(0,2*8),nrow=2) #under assumption of independence
for (i in 1:(8)){
  estimates_n_sum3[1,i]<-fitdistr(PV13h[,i],"normal")$estimate[1] #estimate of mean 
  estimates_n_sum3[2,i]<-fitdistr(PV13h[,i],"normal")$estimate[2] #estimate of std er
  }
```


For every interval of 4h, let $W$ be the random variable distributed according to a univariate normal distribution. 

```{r}
PV14h<-matrix(rep(0,(6*372)),nrow=372) #4h values -> take means
for(i in 1:372){
  for (j in 1:(6)){
    PV14h[i,j]<-mean(PV1h[i,((j-1)*4+1):(j*4)])
  }
}
estimates_n_sum4<-matrix(rep(0,2*6),nrow=2) #under assumption of independence
for (i in 1:(6)){
  estimates_n_sum4[1,i]<-fitdistr(PV14h[,i],"normal")$estimate[1] #estimate of mean 
  estimates_n_sum4[2,i]<-fitdistr(PV14h[,i],"normal")$estimate[2] #estimate of std er
  }
```

Now let's compute mean and standard error of the distribution of $\frac{1}{4} X+Y+Z+W$:
```{r}
mu<-rep(0,24*4)
std<-rep(0,24*4)
for(i in 1:6){
  mu[((i-1)*4*4):(i*4*4)]<-mu[((i-1)*4*4):(i*4*4)]+estimates_n_sum4[1,i]
  std[((i-1)*4*4):(i*4*4)]<-std[((i-1)*4*4):(i*4*4)]+estimates_n_sum4[2,i]
  }
for(i in 1:8){
  mu[((i-1)*3*4):(i*3*4)]<-mu[((i-1)*3*4):(i*3*4)]+estimates_n_sum3[1,i]
  std[((i-1)*3*4):(i*3*4)]<-std[((i-1)*3*4):(i*3*4)]+estimates_n_sum3[2,i]
  }
for(i in 1:24){
  mu[((i-1)*4):(i*4)]<-mu[((i-1)*4):(i*4)]+estimates_n_sum2[1,i]
  std[((i-1)*4):(i*4)]<-std[((i-1)*4):(i*4)]+estimates_n_sum2[2,i]
  }
for(i in 1:24*4){
  mu[i]<-mu[i]+estimates_n_ind[1,i]
  std[i]<-std[i]+estimates_n_ind[2,i]
  }
  
mu<-mu*1/4
std<-std*1/4

```


Let's visualize the expected value we estimated: 

```{r}

plot(mu,xlab="time [h]", ylab="radiation", type="l", col="red", ylim=c(0,3000))
legend(70,3000, c("exp value"), col=c("red"), lty=c(1))
```


##Sample generation

###Sample generation - dependent for 1h intervals\

```{r, include=FALSE,cache=FALSE}
#Just ignore the following lines!
#
# sigma<-estimates_n_dep$sigmahat #24x24 matrix
# mu<-estimates_n_dep$muhat
# A<-chol(sigma)
# 
# N<-10000 #sample size
# 
# rn<-rep(0,24)
# Nstdsample<-matrix(rep(0,24*N),ncol=24) #matrix to store std norm sample
# for (i in 1:N){
#   rn<-rnorm(24)
#   Nstdsample[i,]<-rn
# }
# 
# Nsample<-matrix(rep(0,24*N),ncol=24)#matrix to store N(mu,sigma) sample
# for(i in 1:N){
#   Rea<-rep(0,24)
#   Nstd<-Nstdsample[i,]
#   Rea<-t(A)%*%Nstd+mu
#   Nsample[i,]<-Rea
# }
```

We generate a sample of size $N$ of a **multivariate** normal distribution with the parameters estimated above: 

```{r}
N<-1000
Nsample<-mvrnorm(n=N, estimates_n_dep$muhat, estimates_n_dep$sigmahat)
```



```{r, include=FALSE,cache=FALSE}
#To validate this sample generation, we estimate its distribution and standard error and compare them with the parameters we used:
test<-mlest(Nsample)
diffmu<-estimates_n_dep$muhat-test$muhat
diffsigma<-estimates_n_dep$sigmahat-test$sigmahat
max(diffmu)
max(diffsigma)
#These values show us that a sample size of $N=1000$ is probably not enough. \
```



To visualize, the first three realizations that were generated, look like this: 
```{r}
plot(Nsample[1,], type="l", xlim=c(4,21), ylab="radiation", xlab="time [h]", ylim=c(0,3500))
lines(Nsample[2,], type="l", col="red")
lines(Nsample[3,], type="l", col="blue")
lines(Nsample[4,], type="l", col="green")
lines(Nsample[5,], type="l", col="pink")
lines(Nsample[6,], type="l", col="orange")

```


###Sample generation - independent for 15min intervals


We generate a sample of $N$ **independently** identically distributed random variables of the normal distribution with the parameters estimated above: 

```{r}
N<-10000
N_ind_sample<-matrix(rep(0,24*4*N), ncol=24*4) # 1 column for 1 time interval, 1 row for 1 realization
for(i in 1:N){
N_ind_sample[i,]<-rnorm(24*4,estimates_n_ind[1,], estimates_n_ind[2,])
}
```

To visualize, the first three realizations that were generated, look like this: 
```{r}
plot(N_ind_sample[1,], type="l", xlim=c(25,75), ylab="radiation", xlab="time [15min]")
lines(N_ind_sample[2,], type="l", col="red")
lines(N_ind_sample[3,], type="l", col="blue")

```



#Weibull distribution\

##Parameter estimation\

###Parameter estimation - independent for 15min intervals\

We assume **independently** distributed values for every 15 minutes and estimate $a$ and $b$ called the *shape parameter* and the *scale parameter*.

```{r}
estimates_w_ind<-matrix(rep(0,2*96),nrow=2) #under assumption of independence
for (i in 1:(4*24)){
  coli<-PV1quh[,i]
  coli<-coli[coli>0.001]
  if(length(coli)>0){
  estimates_w_ind[1,i]<-fitdist(coli,"weibull", lower=c(0.001,0.001))$estimate[1] #estimate of shape paramter
  estimates_w_ind[2,i]<-fitdist(coli,"weibull", lower=c(0.001,0.001))$estimate[2] #estimate of scale parameter
  } 
  else {
    estimates_w_ind[1,i]<-0
    estimates_w_ind[2,i]<-0
  }
  }
estimates_w_ind
```

##Sample generation

###Sample generation - independent for 15min intervals\

```{r}
N<-10000
W_ind_sample<-matrix(rep(0,24*4*N), ncol=24*4) # 1 column for 1 time interval, 1 row for 1 realization
for(i in 1:N){
W_ind_sample[i,]<-rweibull(24*4,estimates_w_ind[1,], estimates_w_ind[2,])
}
```

To visualize, the first three realizations that were generated, look like this: 

```{r}
plot(W_ind_sample[1,], type="l", xlim=c(25,75), ylab="radiation", xlab="time [15min]")
lines(W_ind_sample[2,], type="l", col="red")
lines(W_ind_sample[3,], type="l", col="blue")

```

```{r}
plot(PV1quh[10,], type="l", xlim=c(25,75), ylab="radiation", xlab="time [15min]")
lines(PV1quh[88,], type="l", col="red")
lines(PV1quh[9,], type="l", col="blue")

```



